{\bf {\Large Parameter sensitivity and identifiability for a biogeochemical model of hypoxia in the northern {G}ulf of {M}exico}}

### 2016-08-23 ident_manu.Rnw
sensalltile redone for p1z1

%knitr options
<<setup,include=F,cache=F>>=
library(knitr)
# set global chunk options
opts_chunk$set(fig.path = 'figs/', fig.align = 'center', fig.show = 'hold', message = F, echo = F, results = 'asis', dev = 'pdf', dev.args = list(family = 'serif'), fig.pos = '!ht', warning = F)
options(replace.assign = TRUE, width = 90)

source('R/funcs.r')
library(WRTDStidal)
library(tidyr)
library(ggplot2)
library(dplyr)
library(gridExtra)
library(Hmisc)
library(ggrepel)
library(english)
library(purrr)
@

@@

@@

Results from FishTank are based on time-dependent differential equations that describe energy flow between phytoplankton and zooplankton groups as affected by nutrient uptake rates, organic matter inputs and losses, inherent optical properties, sediment diagenesis, and temperature (\citealt{Penta08,Eldridge10}, see appendix in \citetalias{LehrterIR}).  A total of 108 equations are estimated at each time step to return a value for each of the 36 state variables described by the model.  In addition to the initial conditions, 251 parameter values for each of the equations are also supplied at model execution.  These parameters define relationships among fixed effects in the equations and represent ecological properties described by the model that influence hypoxia formation.  Values for each of the parameters were based on estimates from the literature, field or laboratory-based measurements, or expert knowledge in absence of the former.  As such, a sensitivity analysis of parameter values is warranted given that, for example, literature or field-based estimates may not apply under all scenarios or expert knowledge is not completely certain \citep{Refsgaard07}.  The sensitivity of \ac{do} to perturbations of all relevant parameters for the 108 equations was estimated from January 1\textsuperscript{st} to December 31\textsuperscript{st}, 2006 by running FishTank at a timestep of five minutes.  Initial screening removed 73 irrelevant parameters that were not relevant such that 178 were retained for the sensitivity analysis.  Irrelevant parameters were removed for several reasons, e.g., a parameter was only relevant for the hydrodynamic model, was considered a physical constant, or had no effect given initial conditions.  For simplicity, the remaining parameters were grouped into one of six categories based on their respective equations: optics ($n = $ \Sexpr{nparms$Optics} parameters), organic matter (\Sexpr{nparms$`Organic Matter`}), phytoplankton (\Sexpr{nparms$Phytoplankton}), temperature (\Sexpr{nparms$Temperature}), and zooplankton (\Sexpr{nparms$Zooplankton}).  A full description is available as an appendix in \citepalias{LehrterIR}.

@@

\noindent where the estimate depended on the change in the predicted value for response variable $y$ divided by the change in the parameter $\Theta_j$ multiplied by the quotient of scaling factors $w$ for each.  The scaling factors, $w_{\Theta_j}$ for the parameter $\Theta_j$ and $w_{y_i}$ for response variable $y_i$, were set as the default value of the unperturbed parameter and the predicted value of $y_i$ after perturbation \citep{Soetaert10}.  The scaling ensures the estimates are unitless such that the relative magnitudes provide a comparison for model sensitivity to parameter changes that may vary in scale.  Estimates for $S_{ij}$ were summarized as $L1$ and $L2$ across the time series to obtain individual sensitivity values of \ac{do} in response to a change in parame ter $j$:

@@

% for inline
<<echo = FALSE, cache = F>>=
data(sens_ests)
data(sens_ests_cat)

# number of parameters in each category
bycat <- table(sens_ests_cat$Category)
nms <- names(bycat)
bycat <- as.numeric(bycat)
names(bycat) <- nms
bycat <- toeng(bycat)

# percent of parameters in each category
lncats <- table(sens_ests_cat$Category) %>% 
  `/` (table(sort(parcats2(as_df = TRUE)$cats))) %>% 
  `*` (100) %>% 
  form_fun(nsm_val = 0) %>% 
  as.list %>% 
  lapply(., paste0, '\\%')

# L1 range
sensrng <- with(sens_ests_cat, which(error %in% range(error))) %>% 
  sens_ests_cat[., ] %>% 
  mutate(
    error = sapply(error, scinot),
    Parameter = par_txt(Parameter)
    )

# L1 range within categories
sensrngcat <- split(sens_ests_cat, sens_ests_cat$Category) %>% 
  lapply(., function(x){

    with(x, which(error %in% range(error))) %>% 
      x[., ] %>% 
      mutate(
        Parameter = par_txt(Parameter), 
        error = sapply(error, scinot)
      )

    }
  )

# average L1, all and by category
sensave <- scinot(mean(sens_ests_cat$error))
sensave_cat <- split(sens_ests_cat, sens_ests_cat$Category) %>% 
  lapply(., function(x) scinot(mean(x$error)))

# phyto errors
phytogrps <- filter(sens_ests_cat, Category == 'Phytoplankton') %>% 
  separate(Parameter, c('parm', 'grp'), sep = '_') %>% 
  group_by(grp) %>%
  summarise(grpave = mean(error)) %>% 
  arrange(-grpave) %>% 
  mutate(grpave = sapply(grpave, scinot))

# zoop errors 
zoopgrps <- filter(sens_ests_cat, Category == 'Zooplankton') %>% 
  separate(Parameter, c('parm', 'grp'), sep = '_') %>% 
  group_by(grp) %>%
  summarise(grpave = mean(error)) %>% 
  arrange(-grpave) %>% 
  mutate(grpave = sapply(grpave, scinot))
@

Local sensitivity analyses showed that \ac{do} was sensitive to perturbations in \Sexpr{nrow(sens_ests_cat)} of 178 parameters (\Sexpr{round(100 * nrow(sens_ests_cat)/178, 0)}\% of total) in FishTank (see \cref{fig:sensalltile} for other state variables). Within each parameter category, \ac{do} was sensitive to \Sexpr{bycat$Optics} parameters for optics (\Sexpr{lncats$Optics} of all optic parameters, \cref{tab:optsens}), \Sexpr{bycat$`Organic Matter`} for organic matter (\Sexpr{lncats$`Organic Matter`}, \cref{tab:omsens}), \Sexpr{bycat$Phytoplankton} for phytoplankton (\Sexpr{lncats$Phytoplankton}, \cref{tab:phytosens}), \Sexpr{bycat$Temperature} for temperature (\Sexpr{lncats$Temperature}, \cref{tab:tempsens}), and \Sexpr{bycat$Zooplankton} for zooplankton (\Sexpr{lncats$Zooplankton}, \cref{tab:zoopsens}). Although \ac{do} had the greatest sensitivity to parameters in the zooplankton category (as percentage of total), the relative effects varied. Among all parameters, sensitivity values ranged from $L1 = $ \Sexpr{sensrng$error[1]} for \Sexpr{sensrng$Parameter[1]} (\Sexpr{tolower(sensrng$Category[1])}) to \Sexpr{sensrng$error[2]} for \Sexpr{sensrng$Parameter[2]} (\Sexpr{tolower(sensrng$Category[2])}), whereas average sensitivity among all parameters was $L1 = $ \Sexpr{sensave}.  Within categories, sensitivity ranged from \Sexpr{sensrngcat$Optics$error[1]} (\Sexpr{sensrngcat$Optics$Parameter[1]}) to \Sexpr{sensrngcat$Optics$error[2]} (\Sexpr{sensrngcat$Optics$Parameter[2]}) for optics, \Sexpr{sensrngcat$`Organic Matter`$error[1]} (\Sexpr{sensrngcat$`Organic Matter`$Parameter[1]}) to \Sexpr{sensrngcat$`Organic Matter`$error[2]} (\Sexpr{sensrngcat$`Organic Matter`$Parameter[2]}) for organic matter, \Sexpr{sensrngcat$Phytoplankton$error[1]} (\Sexpr{sensrngcat$Phytoplankton$Parameter[1]}) to \Sexpr{sensrngcat$Phytoplankton$error[2]} (\Sexpr{sensrngcat$Phytoplankton$Parameter[2]}) for phytopankton, \Sexpr{sensrngcat$Temperature$error[1]} (\Sexpr{sensrngcat$Temperature$Parameter[1]}) to \Sexpr{sensrngcat$Temperature$error[2]} (\Sexpr{sensrngcat$Temperature$Parameter[2]}) for temperature, and \Sexpr{sensrngcat$Zooplankton$error[1]} (\Sexpr{sensrngcat$Zooplankton$Parameter[1]}) to \Sexpr{sensrngcat$Zooplankton$error[2]} (\Sexpr{sensrngcat$Zooplankton$Parameter[2]}) for zooplankton (\cref{fig:sensplo}, bottom).  Average sensitivity values in each category were $L1 = $ \Sexpr{sensave_cat$Optics} for optics, \Sexpr{sensave_cat$`Organic Matter`} for organic matter, \Sexpr{sensave_cat$Phytoplankton} for phytoplankton, \Sexpr{sensave_cat$Temperature} for temperature, and \Sexpr{sensave_cat$Zooplankton} for zooplankton.  Within the six phytoplankton groups, \ac{do} was sensitive to the same parameters within each group although the sensitivity magnitudes varied. Average sensitivity across parameters in each phytoplankton groups showed that \ac{do} was most sensitive to the first and third phytoplankton groups (average $L1 = $ \Sexpr{phytogrps[1, 2]}, \Sexpr{phytogrps[2, 2]}), whereas sensitivity to parameters in the remaining groups was much lower (all with average $L1 < 1$). Sensitivity of \ac{do} did not vary considerably between parameters in the two zooplankton groups (average $L1 = $ \Sexpr{zoopgrps[1, 2]}, \Sexpr{zoopgrps[2, 2]} for groups one and two).

Response of \ac{do} to parameter perturbations was not uniform across the time series.  \Cref{fig:sensplo} shows variation for the top parameters within each category.  Because FishTank does not include a spatial component, the estimated \ac{do} trend describes a closed, heterotrophic system where respiration processes eventually remove all \ac{do} from the model space.  The initial decrease in the time series reflects change from the initial conditions outside of the growing seasons (i.e., January, February), the spring/summer increase represents production associated with expected seasonal growth, and the remaining time series from August to the end of year shows complete removal of \ac{do} as respiration processes dominate metabolic activity. Although this time series is not a realistic depiction of an actual system, the biogeochemical model behaves as expected in the absence of the hydrodynamic model.  Accordingly, the interpretation of sensitivity results from the simple model has relevance in an ecological context.  As expected, parameter perturbations had the largest effect during the summer months, although the effects varied.  An inrease in 50\% from the parameter default values generally caused a reduction in \ac{do} during the summer, with the exception of the zooplankton parameter, \Sexpr{sensrngcat$Zooplankton$Parameter[2]}, which caused an increase in \ac{do}.  The phytoplankton parameter, \Sexpr{sensrngcat$Phytoplankton$Parameter[2]}, had the largest effect such that the \ac{do} time series was similar to the default output in April/May, whereas a dramatic decrease was observed in the remaining months.  The effects of perturbations early in the time series (January, February) showed similar patterns such that a reduction in \ac{do} was most common, particularly for the optics (\Sexpr{sensrngcat$Optics$Parameter[2]}) and phytoplankton (\Sexpr{sensrngcat$Phytoplankton$Parameter[2]}) parameters.

\subsection{Subset identifiability}

<<>>=
data(cat_ident)
data(top_ident)

# percent combos less than gamma 15, 10 by category
collcatpers <- lapply(cat_ident, function(x){

  colls <- x$coll
  l15 <- scinot(100 * sum(colls < 15)/length(colls), digits = 1)
  l10 <- scinot(100 * sum(colls < 10)/length(colls), digits = 1)
  c(l15, l10)

  })

# percent combos less than gamma 15, 10 by top in each category
colltoppers <- lapply(top_ident, function(x){

  colls <- x$coll
  l15 <- scinot(100 * sum(colls < 15)/length(colls), digits = 1)
  l10 <- scinot(100 * sum(colls < 10)/length(colls), digits = 1)
  c(l15, l10)

  })

maxtempcoll <- scinot(max(cat_ident$Temperature$coll))
@

The identifiability analyses suggested that most parameter subsets exceeded the thresholds of $\gamma = 10, 15$, providing further justification for using selection heuristics for parameter optimization.  Parameter identifiability (as $\gamma$) increased at different rates depending on the parameter category or the number of top parameters that were selected (\cref{fig:identbox}).  By category, identifiability was lowest for parameter subsets in the phytoplankton (\Sexpr{collcatpers$Phytoplankton[1]}\% less than $\gamma = 15$, \Sexpr{collcatpers$Phytoplankton[2]}\% less than $\gamma = 10$)  and zooplankton categories (\Sexpr{collcatpers$Zooplankton[1]}\%, less than $\gamma = 15$, \Sexpr{collcatpers$Zooplankton[2]}\%, less than $\gamma = 10$), whereas a majority of combinations for temperature were identifiable (\Sexpr{collcatpers$Temperature[1]}\% less than $\gamma = 15$, \Sexpr{collcatpers$Temperature[2]}\%, less than $\gamma = 10$).  All subset combinations for optics and organic matter parameters had $\gamma < 10$.  All parameter subsets for choosing the top, top two, and top three parameters in each category were identifiable (\cref{fig:identbox}), whereas a majority were identifiable for choosing the top four (\Sexpr{colltoppers$top4[1]}\% less than $\gamma = 15$, \Sexpr{colltoppers$top4[2]}\% less than $\gamma = 10$) and top five (\Sexpr{colltoppers$top4[1]}\% less than $\gamma = 15$, \Sexpr{colltoppers$top4[2]}\%, less than $\gamma = 10$) parameters.

<<>>=
temp_coll <- par_txt(c('Tref(nospA+nospZ)_1', 'Tref(nospA+nospZ)_2', 'Tref(nospA+nospZ)_3', 'Tref(nospA+nospZ)_4', 'Tref(nospA+nospZ)_5', 'Tref(nospA+nospZ)_6', 'Tref(nospA+nospZ)_7', 'Tref(nospA+nospZ)_8'))
@

A comparison of average and median identifiability by parameter category and top parameters in each category suggested that individual parameters had large effects on $\gamma$ (\cref{fig:identbox}).  For example, a consistent increase in average $\gamma$ from 2 to 7 parameters in a combination for temperature was observed, whereas median identiability remained low until 6 parameter combinations were evaluated.  Further evaluation showed that identifiability was greatly affected by the inclusion of one or two specific parameters in a subset combination.  \Cref{fig:excltemp} shows the temperature collinearity ($\gamma$) ranges in detail for the parameter subsets in \cref{fig:identbox} before and after excluding individual parameters.  Collinearity increases with more parameters included in a subset, although the increase varies depending on the specific parameter.  Exclusion of the parameters \Sexpr{temp_coll[2]} and \Sexpr{temp_coll[5]} showed that $\gamma$ remained well below the 10, 15 threshold for all parameter combinations.  Morever, inclusion of \Sexpr{temp_coll[1]}, \Sexpr{temp_coll[4]}, \Sexpr{temp_coll[7]}, and \Sexpr{temp_coll[8]} generally reduced collinearity relative to when the parameters were excluded.  Similar analyses identified parameters in other categories that had disproportionate effects on identifiability if included in a subset (see supplementary information).

<<>>=
data(p1z1cat_ident)
data(p1z1top_ident)

# percent combos less than gamma 15, 10 by category
p1z1collcatpers <- lapply(p1z1cat_ident, function(x){

  colls <- x$coll
  l15 <- scinot(100 * sum(colls < 15)/length(colls), digits = 1)
  l10 <- scinot(100 * sum(colls < 10)/length(colls), digits = 1)
  c(l15, l10)

  })

# percent combos less than gamma 15, 10 by top in each category
p1z1colltoppers <- lapply(p1z1top_ident, function(x){

  colls <- x$coll
  l15 <- scinot(100 * sum(colls < 15)/length(colls), digits = 1)
  l10 <- scinot(100 * sum(colls < 10)/length(colls), digits = 1)
  c(l15, l10)

  })
@

Comparison of identifiability between categories showed that phytoplankton and zooplankton had the least identifiable parameter subsets.  As noted above, FishTank includes six phytoplankton and two zooplankton groups to characterize community structure and foodweb dynamics that likely have an important role in hypoxia development.  However, structural equations for each group do not vary considerably such that variation in parameter values primarily control differences between the groups, e.g., large-bodied vs small-bodied plankton, slow-growing vs. fast-growing plankton.  To obtain identifiability estimates of the plankton categories that were independent of groups, the identifiability analyses were re-evaluated using only one phytoplankton and one zooplankton group (\cref{fig:p1z1identbox}). Compared to all groups, evaluating a single group improved identifiability such that a majority of parameter combinations were below the threshold (\Sexpr{p1z1collcatpers$Phytoplankton[1]}\% less than $\gamma = 15$, \Sexpr{p1z1collcatpers$Phytoplankton[2]}\% less than $\gamma = 10$ for Phytoplankton; \Sexpr{p1z1collcatpers$Zooplankton[1]}\%, less than $\gamma = 15$, \Sexpr{p1z1collcatpers$Zooplankton[2]}\%, less than $\gamma = 10$ for zooplankton).  Analysis of identifiability for top parameters in all categories did not show similar changes in identifiability.  Although all combinations for the top one, two, and three parameters in each category were still well below the threshold, selecting the top four and five parameters showed an increase (\Sexpr{p1z1colltoppers$top4[1]}\% less than $\gamma = 15$, \Sexpr{p1z1colltoppers$top4[2]}\% less than $\gamma = 10$) and a decrease (\Sexpr{p1z1colltoppers$top5[1]}\% less than $\gamma = 15$, \Sexpr{p1z1colltoppers$top5[2]}\% less than $\gamma = 10$) in identifiability, respectively (\cref{fig:p1z1identbox}), compared to results that included all phytoplankton and zooplankton groups (\cref{fig:identbox}).

\subsection{Parameter selection}

<<>>=
data(tocal)

>>>>

%knitr options
<<setup, include = F, cache = F>>=
library(knitr)
# set global chunk options
opts_chunk$set(fig.path = 'figs/', fig.align = 'center', fig.show = 'hold', message = F, echo = F, results = 'asis', dev = 'pdf', dev.args = list(family = 'serif'), fig.pos = '!ht', warning = F)
options(replace.assign = TRUE, width = 90)
@

% R dependencies
<<include = F, cache = F>>=
source('R/funcs.r')
library(WRTDStidal)
library(tidyr)
library(ggplot2)
library(dplyr)
library(gridExtra)
library(Hmisc)
library(ggrepel)
library(english)
library(purrr)
@

@@

Results from FishTank are based on time-dependent differential equations that describe energy flow between phytoplankton and zooplankton groups as affected by nutrient uptake rates, organic matter inputs and losses, inherent optical properties, sediment diagenesis, and temperature (\citealt{Penta08,Eldridge10}, see appendix in \citetalias{LehrterIR}).  A total of 108 equations are estimated at each time step to return a value for each of the 36 state variables described by the model.  In addition to the initial conditions, 251 parameter values for each of the equations are also supplied at model execution.  These parameters define relationships among fixed effects in the equations and represent ecological properties described by the model that influence hypoxia formation.  Values for each of the parameters were based on estimates from the literature, field or laboratory-based measurements, or expert knowledge in absence of the former.  As such, a sensitivity analysis of parameter values is warranted given that, for example, literature or field-based estimates may not apply under all scenarios or expert knowledge is not completely certain \citep{Refsgaard07}.  The sensitivity of \ac{do} to perturbations of all relevant parameters for the 108 equations was estimated from January 1\textsuperscript{st} to December 31\textsuperscript{st}, 2006 by running FishTank at a timestep of five minutes.  Initial screening removed 73 irrelevant parameters that were removed for several reasons, e.g., a parameter was only relevant for the hydrodynamic model, was considered a physical constant, or had no effect given initial conditions.

@@

As noted above, FishTank includes six phytoplankton and two zooplankton groups to characterize greater complexity in community structure and foodweb dynamics.  However, structural equations for each group do not vary considerably such that variation in parameter values primarily control differences between the groups, e.g., large-bodied or small-bodied plankton, slow-growing or fast-growing plankton, etc..  Initial analyses indicated that parameter sensitivity of dissolved oxygen was identical within the six phytoplankton and zooplankton groups.  To remove obvious redundancies in the model, the sensitivity analyes were conducted using only one phytoplankton and one zooplankton group.  This further reduced the parameter set....  For simplicity, the remaining parameters were grouped in  to one of six categories based on their respective equations: optics ($n = $ \Sexpr{nparms$Optics} parameters), organic matter (\Sexpr{nparms$`Organic Matter`}), phytoplankton (\Sexpr{nparms$Phytoplankton}), temperature (\Sexpr{nparms$Temperature}), and zooplankton (\Sexpr{nparms$Zooplankton}).  A full description is available as an appendix in \citepalias{LehrterIR}.

@@

\noindent where the estimate depended on the change in the predicted value for response variable $y$ divided by the change in the parameter $\Theta_j$ multiplied by the quotient of scaling factors $w$ for each.  The scaling factors, $w_{\Theta_j}$ for the parameter $\Theta_j$ and $w_{y_i}$ for response variable $y_i$, were set as the default value of the unperturbed parameter and the initial value of $y$ at time $= 0$ \citep{Soetaert10}.  The scaling ensures the estimates are unitless such that the relative magnitudes provide a comparison for model sensitivity to parameter changes and state variables that differ in scale.  Estimates for $S_{ij}$ were summarized as $L1$ and $L2$ across the time series to obtain individual sensitivity values of \ac{do} in response to a change in parame ter $j$:

@@

%' 
%' % for inline
%' <<echo = FALSE, eavl = F, cache = F>>=
%' data(sens_ests)
%' data(sens_ests_cat)
%' 
%' # number of parameters in each category
%' bycat <- table(sens_ests_cat$Category)
%' nms <- names(bycat)
%' bycat <- as.numeric(bycat)
%' names(bycat) <- nms
%' bycat <- toeng(bycat)
%' 
%' # percent of parameters in each category
%' lncats <- table(sens_ests_cat$Category) %>% 
%'   `/` (table(sort(parcats2(as_df = TRUE)$cats))) %>% 
%'   `*` (100) %>% 
%'   form_fun(nsm_val = 0) %>% 
%'   as.list %>% 
%'   lapply(., paste0, '\\%')
%' 
%' # L1 range
%' sensrng <- with(sens_ests_cat, which(error %in% range(error))) %>% 
%'   sens_ests_cat[., ] %>% 
%'   mutate(
%'     error = sapply(error, scinot),
%'     Parameter = par_txt(Parameter)
%'     )
%' 
%' # L1 range within categories
%' sensrngcat <- split(sens_ests_cat, sens_ests_cat$Category) %>% 
%'   lapply(., function(x){
%'     
%'     with(x, which(error %in% range(error))) %>% 
%'       x[., ] %>% 
%'       mutate(
%'         Parameter = par_txt(Parameter), 
%'         error = sapply(error, scinot)
%'       )
%'     
%'     }
%'   )
%' 
%' # average L1, all and by category
%' sensave <- scinot(mean(sens_ests_cat$error))
%' sensave_cat <- split(sens_ests_cat, sens_ests_cat$Category) %>% 
%'   lapply(., function(x) scinot(mean(x$error)))
%'     
%' # phyto errors
%' phytogrps <- filter(sens_ests_cat, Category == 'Phytoplankton') %>% 
%'   separate(Parameter, c('parm', 'grp'), sep = '_') %>% 
%'   group_by(grp) %>%
%'   summarise(grpave = mean(error)) %>% 
%'   arrange(-grpave) %>% 
%'   mutate(grpave = sapply(grpave, scinot))
%' 
%' # zoop errors 
%' zoopgrps <- filter(sens_ests_cat, Category == 'Zooplankton') %>% 
%'   separate(Parameter, c('parm', 'grp'), sep = '_') %>% 
%'   group_by(grp) %>%
%'   summarise(grpave = mean(error)) %>% 
%'   arrange(-grpave) %>% 
%'   mutate(grpave = sapply(grpave, scinot))
%' @
%' 
%' \subsection{Local sensitivity analysis}
%' 
%' Local sensitivity analyses showed that \ac{do} was sensitive to perturbations in \Sexpr{nrow(sens_ests_cat)} of 178 parameters (\Sexpr{round(100 * nrow(sens_ests_cat)/178, 0)}\% of total) in FishTank (see \cref{fig:sensalltile} for other state variables). Within each parameter category, \ac{do} was sensitive to \Sexpr{bycat$Optics} parameters for optics (\Sexpr{lncats$Optics} of all optic parameters, \cref{tab:optsens}), \Sexpr{bycat$`Organic Matter`} for organic matter (\Sexpr{lncats$`Organic Matter`}, \cref{tab:omsens}), \Sexpr{bycat$Phytoplankton} for phytoplankton (\Sexpr{lncats$Phytoplankton}, \cref{tab:phytosens}), \Sexpr{bycat$Temperature} for temperature (\Sexpr{lncats$Temperature}, \cref{tab:tempsens}), and \Sexpr{bycat$Zooplankton} for zooplankton (\Sexpr{lncats$Zooplankton}, \cref{tab:zoopsens}). Although \ac{do} had the greatest sensitivity to parameters in the zooplankton category (as percentage of total), the relative effects varied. Among all parameters, sensitivity values ranged from $L1 = $ \Sexpr{sensrng$error[1]} for \Sexpr{sensrng$Parameter[1]} (\Sexpr{tolower(sensrng$Category[1])}) to \Sexpr{sensrng$error[2]} for \Sexpr{sensrng$Parameter[2]} (\Sexpr{tolower(sensrng$Category[2])}), whereas average sensitivity among all parameters was $L1 = $ \Sexpr{sensave}.  Within categories, sensitivity ranged from \Sexpr{sensrngcat$Optics$error[1]} (\Sexpr{sensrngcat$Optics$Parameter[1]}) to \Sexpr{sensrngcat$Optics$error[2]} (\Sexpr{sensrngcat$Optics$Parameter[2]}) for optics, \Sexpr{sensrngcat$`Organic Matter`$error[1]} (\Sexpr{sensrngcat$`Organic Matter`$Parameter[1]}) to \Sexpr{sensrngcat$`Organic Matter`$error[2]} (\Sexpr{sensrngcat$`Organic Matter`$Parameter[2]}) for organic matter, \Sexpr{sensrngcat$Phytoplankton$error[1]} (\Sexpr{sensrngcat$Phytoplankton$Parameter[1]}) to \Sexpr{sensrngcat$Phytoplankton$error[2]} (\Sexpr{sensrngcat$Phytoplankton$Parameter[2]}) for phytopankton, \Sexpr{sensrngcat$Temperature$error[1]} (\Sexpr{sensrngcat$Temperature$Parameter[1]}) to \Sexpr{sensrngcat$Temperature$error[2]} (\Sexpr{sensrngcat$Temperature$Parameter[2]}) for temperature, and \Sexpr{sensrngcat$Zooplankton$error[1]} (\Sexpr{sensrngcat$Zooplankton$Parameter[1]}) to \Sexpr{sensrngcat$Zooplankton$error[2]} (\Sexpr{sensrngcat$Zooplankton$Parameter[2]}) for zooplankton (\cref{fig:sensplo}, bottom).  Average sensitivity values in each category were $L1 = $ \Sexpr{sensave_cat$Optics} for optics, \Sexpr{sensave_cat$`Organic Matter`} for organic matter, \Sexpr{sensave_cat$Phytoplankton} for phytoplankton, \Sexpr{sensave_cat$Temperature} for temperature, and \Sexpr{sensave_cat$Zooplankton} for zooplankton.  Within the six phytoplankton groups, \ac{do} was sensitive to the same parameters within each group although the sensitivity magnitudes varied. Average sensitivity across parameters in each phytoplankton groups showed that \ac{do} was most sensitive to the first and third phytoplankton groups (average $L1 = $ \Sexpr{phytogrps[1, 2]}, \Sexpr{phytogrps[2, 2]}), whereas sensitivity to parameters in the remaining groups was much lower (all with average $L1 < 1$). Sensitivity of \ac{do} did not vary considerably between parameters in the two zooplankton groups (average $L1 = $ \Sexpr{zoopgrps[1, 2]}, \Sexpr{zoopgrps[2, 2]} for groups one and two).
%' 
%' Response of \ac{do} to parameter perturbations was not uniform across the time series.  \Cref{fig:sensplo} shows variation for the top parameters within each category.  Because FishTank does not include a spatial component, the estimated \ac{do} trend describes a closed, heterotrophic system where respiration processes eventually remove all \ac{do} from the model space.  The initial decrease in the time series reflects change from the initial conditions outside of the growing seasons (i.e., January, February), the spring/summer increase represents production associated with expected seasonal growth, and the remaining time series from August to the end of year shows complete removal of \ac{do} as respiration processes dominate metabolic activity. Although this time series is not a realistic depiction of an actual system, the biogeochemical model behaves as expected in the absence of the hydrodynamic model.  Accordingly, the interpretation of sensitivity results from the simple model has relevance in an ecological context.  As expected, parameter perturbations had the largest effect during the summer months, although the effects varied.  An inrease in 50\% from the parameter default values generally caused a reduction in \ac{do} during the summer, with the exception of the zooplankton parameter, \Sexpr{sensrngcat$Zooplankton$Parameter[2]}, which caused an increase in \ac{do}.  The phytoplankton parameter, \Sexpr{sensrngcat$Phytoplankton$Parameter[2]}, had the largest effect such that the \ac{do} time series was similar to the default output in April/May, whereas a dramatic decrease was observed in the remaining months.  The effects of perturbations early in the time series (January, February) showed similar patterns such that a reduction in \ac{do} was most common, particularly for the optics (\Sexpr{sensrngcat$Optics$Parameter[2]}) and phytoplankton (\Sexpr{sensrngcat$Phytoplankton$Parameter[2]}) parameters.
%' 
%' \subsection{Subset identifiability}
%' 
%' <<eval = F>>=
%' data(cat_ident)
%' data(top_ident)
%' 
%' # percent combos less than gamma 15, 10 by category
%' collcatpers <- lapply(cat_ident, function(x){
%'   
%'   colls <- x$coll
%'   l15 <- scinot(100 * sum(colls < 15)/length(colls), digits = 1)
%'   l10 <- scinot(100 * sum(colls < 10)/length(colls), digits = 1)
%'   c(l15, l10)
%'   
%'   })
%' 
%' # percent combos less than gamma 15, 10 by top in each category
%' colltoppers <- lapply(top_ident, function(x){
%'   
%'   colls <- x$coll
%'   l15 <- scinot(100 * sum(colls < 15)/length(colls), digits = 1)
%'   l10 <- scinot(100 * sum(colls < 10)/length(colls), digits = 1)
%'   c(l15, l10)
%'   
%'   })
%' 
%' maxtempcoll <- scinot(max(cat_ident$Temperature$coll))
%' @
%' 
%' The identifiability analyses suggested that most parameter subsets exceeded the thresholds of $\gamma = 10, 15$, providing further justification for using selection heuristics for parameter optimization.  Parameter identifiability (as $\gamma$) increased at different rates depending on the parameter category or the number of top parameters that were selected (\cref{fig:identbox}).  By category, identifiability was lowest for parameter subsets in the phytoplankton (\Sexpr{collcatpers$Phytoplankton[1]}\% less than $\gamma = 15$, \Sexpr{collcatpers$Phytoplankton[2]}\% less than $\gamma = 10$)  and zooplankton categories (\Sexpr{collcatpers$Zooplankton[1]}\%, less than $\gamma = 15$, \Sexpr{collcatpers$Zooplankton[2]}\%, less than $\gamma = 10$), whereas a majority of combinations for temperature were identifiable (\Sexpr{collcatpers$Temperature[1]}\% less than $\gamma = 15$, \Sexpr{collcatpers$Temperature[2]}\%, less than $\gamma = 10$).  All subset combinations for optics and organic matter parameters had $\gamma < 10$.  All parameter subsets for choosing the top, top two, and top three parameters in each category were identifiable (\cref{fig:identbox}), whereas a majority were identifiable for choosing the top four (\Sexpr{colltoppers$top4[1]}\% less than $\gamma = 15$, \Sexpr{colltoppers$top4[2]}\% less than $\gamma = 10$) and top five (\Sexpr{colltoppers$top4[1]}\% less than $\gamma = 15$, \Sexpr{colltoppers$top4[2]}\%, less than $\gamma = 10$) parameters.  
%' 
%' <<eval = F>>=
%' temp_coll <- par_txt(c('Tref(nospA+nospZ)_1', 'Tref(nospA+nospZ)_2', 'Tref(nospA+nospZ)_3', 'Tref(nospA+nospZ)_4', 'Tref(nospA+nospZ)_5', 'Tref(nospA+nospZ)_6', 'Tref(nospA+nospZ)_7', 'Tref(nospA+nospZ)_8'))
%' @
%' 
%' A comparison of average and median identifiability by parameter category and top parameters in each category suggested that individual parameters had large effects on $\gamma$ (\cref{fig:identbox}).  For example, a consistent increase in average $\gamma$ from 2 to 7 parameters in a combination for temperature was observed, whereas median identiability remained low until 6 parameter combinations were evaluated.  Further evaluation showed that identifiability was greatly affected by the inclusion of one or two specific parameters in a subset combination.  \Cref{fig:excltemp} shows the temperature collinearity ($\gamma$) ranges in detail for the parameter subsets in \cref{fig:identbox} before and after excluding individual parameters.  Collinearity increases with more parameters included in a subset, although the increase varies depending on the specific parameter.  Exclusion of the parameters \Sexpr{temp_coll[2]} and \Sexpr{temp_coll[5]} showed that $\gamma$ remained well below the 10, 15 threshold for all parameter combinations.  Morever, inclusion of \Sexpr{temp_coll[1]}, \Sexpr{temp_coll[4]}, \Sexpr{temp_coll[7]}, and \Sexpr{temp_coll[8]} generally reduced collinearity relative to when the parameters were excluded.  Similar analyses identified parameters in other categories that had disproportionate effects on identifiability if included in a subset (see supplementary information).
%' 
%' <<eval = F>>=
%' data(p1z1cat_ident)
%' data(p1z1top_ident)
%' 
%' # percent combos less than gamma 15, 10 by category
%' p1z1collcatpers <- lapply(p1z1cat_ident, function(x){
%'   
%'   colls <- x$coll
%'   l15 <- scinot(100 * sum(colls < 15)/length(colls), digits = 1)
%'   l10 <- scinot(100 * sum(colls < 10)/length(colls), digits = 1)
%'   c(l15, l10)
%'   
%'   })
%' 
%' # percent combos less than gamma 15, 10 by top in each category
%' p1z1colltoppers <- lapply(p1z1top_ident, function(x){
%'   
%'   colls <- x$coll
%'   l15 <- scinot(100 * sum(colls < 15)/length(colls), digits = 1)
%'   l10 <- scinot(100 * sum(colls < 10)/length(colls), digits = 1)
%'   c(l15, l10)
%'   
%'   })
%' @
%' 
